### 2023-05-25 paper/paperDraft.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  dvipsnames,enabledeprecatedfontcommands]{scrartcl}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
%\documentclass{article}

% %packages
\usepackage{booktabs}
%\usepackage[left]{showlabels}
\usepackage{multirow}
\usepackage{subcaption}

\usepackage{graphicx}
%\usepackage{longtable}
\usepackage{supertabular, booktabs}
\usepackage{ragged2e}
\usepackage{etex}
%\usepackage{yfonts}
\usepackage{marvosym}
\usepackage[notextcomp]{kpfonts}
\usepackage{nicefrac}
\newcommand*{\QED}{\hfill \footnotesize {\sc Q.e.d.}}
\usepackage{floatrow}

\usepackage[textsize=footnotesize]{todonotes}

%\linespread{1.5}
\newcommand{\indep}{\!\perp \!\!\! \perp\!}

\setlength{\parindent}{5mm}
\setlength{\parskip}{1pt}

%language
%\usepackage{times}
\usepackage{mathptmx}
\usepackage[scaled=0.86]{helvet}
\usepackage{t1enc}
%\usepackage[utf8x]{inputenc}
%\usepackage[polish]{babel}
%\usepackage{polski}

%AMS
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathtools}

\usepackage{geometry}
 \geometry{a4paper,left=35mm,top=20mm,}

%environments
\newtheorem{fact}{Fact}

%abbreviations
\newcommand{\ra}{\rangle}
\newcommand{\la}{\langle}
\newcommand{\n}{\neg}
\newcommand{\et}{\wedge}
\newcommand{\jt}{\rightarrow}
\newcommand{\ko}[1]{\forall  #1\,}
\newcommand{\ro}{\leftrightarrow}
\newcommand{\exi}[1]{\exists\, {_{#1}}}
\newcommand{\pr}[1]{\ensuremath{\mathsf{P}(#1)}}
\newcommand{\cost}{\mathsf{cost}}
\newcommand{\benefit}{\mathsf{benefit}}
\newcommand{\ut}{\mathsf{ut}}

\newcommand{\odds}{\mathsf{Odds}}
\newcommand{\ind}{\mathsf{Ind}}
\newcommand{\nf}[2]{\nicefrac{#1\,}{#2}}
\newcommand{\R}[1]{\texttt{#1}}
\newcommand{\prr}[1]{\mbox{$\mathtt{P}_{prior}(#1)$}}
\newcommand{\prp}[1]{\mbox{$\mathtt{P}_{posterior}(#1)$}}

\newtheorem{q}{\color{blue}Question}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}[fact]

%technical intermezzo
%---------------------

\newcommand{\intermezzoa}{
	\begin{minipage}[c]{13cm}
	\begin{center}\rule{10cm}{0.4pt}

	\tiny{\sc Optional Content Starts}

	\vspace{-1mm}

	\rule{10cm}{0.4pt}\end{center}
	\end{minipage}\nopagebreak 
	}

\newcommand{\intermezzob}{\nopagebreak 
	\begin{minipage}[c]{13cm}
	\begin{center}\rule{10cm}{0.4pt}

	\tiny{\sc Optional Content Ends}

	\vspace{-1mm}

	\rule{10cm}{0.4pt}\end{center}
	\end{minipage}
	}

%--------------------

\newtheorem*{reply*}{Reply}
\usepackage{enumitem}
\newcommand{\question}[1]{\begin{enumerate}[resume,leftmargin=0cm,labelsep=0cm,align=left]
\item #1
\end{enumerate}}

\usepackage{float}

% \usepackage{sectsty}
% \sectionfont{\fontsize{12}{11}\selectfont}
% \subsectionfont{\fontsize{11}{11}\selectfont}

% \setbeamertemplate{blocks}[rounded][shadow=true]
% \setbeamertemplate{itemize items}[ball]
% \AtBeginPart{}
% \AtBeginSection{}
% \AtBeginSubsection{}
% \AtBeginSubsubsection{}
% \setlength{\emergencystretch}{0em}
% \setlength{\parskip}{0pt}

\usepackage[authoryear]{natbib}

%\bibliographystyle{apalike}

\usepackage{tikz}
\usetikzlibrary{positioning,shapes,arrows}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Taking uncertainty in word embedding bias estimation seriously: a bayesian approach},
  pdfauthor={Alicja Dobrzeniecka and Rafal Urbaniak},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={blue},
  pdfcreator={LaTeX via pandoc}}

\title{Taking uncertainty in word embedding bias estimation seriously: a
bayesian approach}
\author{Alicja Dobrzeniecka and Rafal Urbaniak}
\date{}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

It has been suggested\footnote{See for instance {[}1,4,5,8,11,12{]}.}
that language models can learn implicit biases that reflect harmful
stereotypical thinking---for example, the (vector corresponding to the)
word \textit{she} might be much closer in the vector space to the word
\textit{cooking} than the word \textit{he}. Such phenomena are
undesirable at least in some downstream tasks, such as web search,
recommendations, and so on. To investigate such issues, several measures
of bias in word embeddings have been formulated and applied. Our goal is
to use two prominent examples of such measures to argue that this
approach oversimplifies the situation and to develop a Bayesian
alternative.

A common approach in natural language processing is to represent words
by vectors of real numbers---such representations are called
\emph{embeddings}. One way to construct an embedding---we will focus our
attention on non-contextual language models\footnote{One example of a
  contextualized representation is BERT. Another is GPT.}---is to use a
large corpus to train a neural network to assign vectors to words in a
way that optimizes for co-occurrence prediction accuracy. Such vectors
can then be compared in terms of their similarity---the usual measure is
cosine similarity---and the results of such comparisons can be used in
downstream tasks. Roughly speaking, cosine similarity is an imperfect
mathematical proxy for semantic similarity {[}14{]}.

One response to the raising of the issue of bias in natural language
models might be to say that there is not much point to reflecting on
such biases, as they are unavoidable. This unavoidability might seem in
line with the arguments to the effect that learning algorithms are
always value-laden {[}9{]}: they employ inductive methods that require
design-, data-, or risk-related decisions that have to be guided by
extra-algorithmic considerations. Such choices necessarily involve value
judgments and have to do, for instance, with what simplifications or
risks one finds acceptable. Admittedly, algorithmic decision making
cannot fulfill the value-free ideal, but this only means that even more
attention needs to be paid to the values underlying different techniques
and decisions, and to the values being pursued in a particular use of an
algorithm.

Another response might be to insist that there is no bias introduced by
the use of machine learning methods here, since the algorithm is simply
learning to correctly predict co-occurrences based on what ``reality''
looks like. However, this objection overlooks the fact that we, humans,
are the ones who construct this linguistic reality, which is shaped in
part by the natural language processing tools we use on a massive scale.
Sure, if there is unfairness and our goal is to diagnose it, we should
do complete justice to learning it in the model used to study it. One
example of this approach is {[}3{]}, where the authors use language
models to study the shape of certain biases across a century.

However, if our goal is to develop downstream tools that perform tasks
that we care about without further perpetuating or exacerbating harmful
stereotypes, we still have good reasons to try to minimize the negative
impact. Moreover, it is often not the case that the corpora mirror
reality---to give a trivial example, heads are spoken of more often than
kidneys, but this does not mean that kidneys occur much less often in
reality than heads. To give a more relevant example, the
disproportionate association of female words with female occupations in
a corpus actually greatly exaggerates the actual lower disproportion in
the real distribution of occupations {[}6{]}.

In what follows, we focus on two popular measures of bias applicable to
many existing word embeddings, such as
\emph{GoogleNews},\footnote{GoogleNews-vectors-negative300, available at  \url{https://github.com/mmihaltz/word2vec-GoogleNews-vectors}.}
\emph{Glove}\footnote{Available at \url{https://nlp.stanford.edu/projects/glove/}.}
and
\emph{Reddit Corpus}\footnote{Reddit-L2 corpus, available at  \url{http://cl.haifa.ac.il/projects/L2/}.}:
\emph{Word Embedding Association Test} (\textsf{WEAT}) {[}8{]}, and
\emph{Mean Average Cosine Distance} (\textsf{MAC}) {[}12{]}. We first
explain how these measures are supposed to work. Then we argue that they
are problematic for various reasons---the key one being that by
pre-averaging data they manufacture false confidence, which we
illustrate in terms of simulations showing that the measures often
suggest the existence of bias even if by design it is non-existent in a
simulated dataset.

We propose to replace them with a Bayesian data analysis, which not only
provides more modest and realistic assessment of the uncertainty
involved, but in which hierarchical models allow for inspection at
various levels of granularity. Once we introduce the method, we apply it
to multiple word embeddings and results of supposed debiasing, putting
forward some general observations that are not exactly in line with the
usual picture painted in terms of \textsf{WEAT} or \textsf{MAC}.

Most of the problems that we point out generalize to any existing
approach that focuses on chasing a single numeric metric of bias. (1)
They treat the results of pre-averaging as raw data in statistical
significance tests, which in this context is bound to overestimate
significance. We show similar results can easily be obtained when
sampling from null models with no bias. (2) The word list sizes and
sample sizes used in the studies are usually small,\footnote{Depending
  on a list for {[}8{]} the range for protected words is between 13 and
  100, and for attributes between 16 and 25; for {[}12{]} the range for
  protected words is between 14 and 18, and for attributes between 11
  and 25.} (3) Many studies do not use any control predicates, such as
random neutral words or neutral human predicates for comparison.

On the constructive side, we develop and deploy our method, and the
results are, roughly, as follows. (A) Posterior density intervals are
fairly wide and the average differences between associated, different
and neutral human predicates, are not very large. (B) A preliminary
inspection suggests that the desirability of changes obtained by the
usual debiasing methods is debatable.

In Section \ref{sec:two} we describe the two key measures discussed in
this paper, \textsf{WEAT} and \textsf{MAC}, explaining how they are
calculated and how they are supposed to work. In Section
\ref{sec:challenges} we first argue in Subsection
\ref{subsec:interpretability}, that it is far from clear how results
given in terms of \textsf{WEAT} or \textsf{MAC} are to be interpreted.
Second, in Subsection \ref{subsec:problems} we explain the statistical
problems that arise when one uses pre-averaged data in such contexts, as
these measures do. In Section \ref{sec:bayesian} we explain the
alternative Bayesian approach that we propose. In Section
\ref{sec:results} we elaborate on the results that it leads to,
including a somewhat skeptical view of the efficiency of debiasing
methods, discussed in Subsection \ref{subsec:rethinking}. Finally, in
Section \ref{sec:related} we spend some time placing our results in the
ongoing discussions.\footnote{\textbf{Disclaimer:} throughout the paper
  we will be mentioning and using word lists and stereotypes we did not
  formulate, which does not mean we condone any judgment made therein or
  underlying a given word selection. For instance, the Gender dataset
  does not recognize non-binary categories, and yet we use it without
  claiming that such categories should be ignored.}

\hypertarget{two-measures-of-bias-weat-and-mac}{%
\section{Two measures of bias: WEAT and
MAC}\label{two-measures-of-bias-weat-and-mac}}

\label{sec:two}

Conceptually, the idea is that if a particular harmful stereotype is
learned in a particular embedding, then certain groups of words will be
systematically closer to (or further from) each other. This gives rise
to the idea of protected groups---for example, in guiding online search
completion or recommendation, female words might require protection in
that they should not be systematically closer to stereotypically female
job names, such as ``nurse'', ``librarian'', ``waitress'', and male
words require protection in that they should not be systematically
closer to toxic masculinity stereotypes, such as ``tough'', ``never
complaining'' or ``macho''.\footnote{However, for some research-related
  purposes, such as the study of stereotypes across history {[}3{]},
  embeddings which do not protect certain classes may also be useful.}

The key role in the measures to be discussed is played by the notion of
cosine distance (or, symmetrically, by cosine similarity). These are
defined as follows:\footnote{Here, ``\(-\)'' stands for point-wise
  difference, ``\(\cdot\)'' stands for the dot product operation, and
  \(\lVert a\rVert = \sqrt{(a \cdot a)}\).}
\(^{\!\!\! , \,}\)\footnote{Note that this terminology is slightly
  misleading, as mathematically cosine distance is not a distance
  measure, because it does not satisfy the triangle inequality, as
  generally
  \(\mathsf{cosineDistance}(A,C) \not \leq \mathsf{cosineDistance}(A,B)+ \mathsf{cosineDistance}(B,C)\).
  We will keep using this mainstream terminology.}
\begin{align} \tag{Sim}
\mathsf{cosineSimilarity}(A,B) & = \frac{A \cdot B}{\lVert  A \rVert \,\lVert B \rVert}
\\
\tag{Distance}
\mathsf{cosineDistance}(A,B) &  = 1 - \mathsf{cosineSimilarity}(A,B).
\end{align}

One of the first measures of bias has been developed in {[}1{]}. The
general idea is that a certain topic is associated with a vector of real
numbers (the topic ``direction''), and the bias of a word is
investigated by considering the projection of its corresponding vector
on this direction. For instance, in {[}1{]}, the gender direction
\(\mathsf{gd}\) is obtained by taking the differences of the vectors
corresponding to ten different gendered pairs (such as
\(\overrightarrow{she} - \overrightarrow{he}\) or
\(\overrightarrow{girl} - \overrightarrow{boy}\)), and then identifying
their principal component.\footnote{Roughly, the principal component is
  the vector obtained by projecting the data points on their linear
  combination in a way that maximizes the variance of the projections.}
The gender bias of a word \(w\) is then understood as \(w\)'s projection
on the gender direction: \(\vec{w} \cdot gd\) (which, after normalizing
by dividing by \(\lVert w \rVert \,\lVert \mathsf{gd} \rVert\), is the
same as cosine similarity). Given a list \(N\) of supposedly gender
neutral words,\footnote{We follow the methodology used in the debate in
  assuming that there is a class of words identified as more or less
  neutral, such as \emph{ballpark, eat, walk, sleep, table}, whose
  average similarity to the gender direction (or other protected words)
  is around 0. We will have more to say about this assumption when we
  describe our dataset construction.} and the gender direction
\(\mathsf{gd}\), the direct gender bias is defined as the average cosine
similarity of the words in \(N\) from \(\mathsf{gd}\) (\(c\) is a
parameter determining how strict we want to be): \begin{align*}
\mathsf{directBias_c(N,gd)} & = \frac{\sum_{w\in N}\vert \mathsf{cos}(\vec{w},\mathsf{gd})\vert^c}{\vert N \vert }
\end{align*} \normalsize

The use of projections in bias estimation has been criticized for
instance in {[}5{]}, where it is pointed out that while a higher average
similarity to the gender direction might be an indicator of bias with
respect to a given class of words, it is only one possible manifestation
of it, and reducing the cosine similarity to such a projection may not
be sufficient to eliminate bias. For instance, ``math'' and ``delicate''
might be equally similar to a pair of opposed explicitly gendered words
(\emph{she}, \emph{he}), while being closer to quite different
stereotypical attribute words (such as \emph{scientific} or
\emph{caring}). Further, it is observed in {[}5{]} that most word pairs
retain similarity under debiasing meant to minimize projection-based
bias.\footnote{In {[}1{]} another method which involves analogies and
  their evaluations by human users on Mechanical Turk is also used. We
  do not discuss this method in this paper, see its criticism in
  {[}16{]}.}

A measure of bias in word embeddings which does not employ similarities
to pre-calculated protected-class-level directions, the Word Embedding
Association Test (\textsf{WEAT}), has been proposed in {[}8{]}. The idea
here is that the bias between two sets of target words, \(X\) and \(Y\)
(we call them protected words), should be quantified in terms of the
cosine similarity between the protected words and attribute words coming
from two sets of stereotype attribute words, \(A\) and \(B\) (we will
call them attributes). For instance, \(X\) might be a set of male names,
\(Y\) a set of female names, \(A\) might contain stereotypically
male-related, and \(B\) stereotypically female-related career words. The
association difference for a particular word \(t\) (belonging to either
\(X\) or \(Y\)) is:

\vspace{-2mm}

\begin{align}
\label{eq:stAB}
\mathsf{s}(t,A,B) & = \frac{\sum_{a\in A}\mathsf{cos}(t,a)}{\vert A\vert} - \frac{\sum_{b\in B}\mathsf{cos}(t,b)}{\vert B\vert}
\end{align} \normalsize \noindent then, the association difference
between \(A\) a \(B\) is:

\begin{align}
\label{eq:sXYAB}
\mathsf{s}(X,Y,A,B) & = \sum_{x\in X} \mathsf{s}(x,A,B) -  \sum_{y\in Y} \mathsf{s}(y,A,B)
\end{align}

\noindent The authors use it as a test statistic in some tests, and the
assumption that \(X\) and \(Y\) are of the same size is necessary for
the statistic to make sense. In the final measure (of effect size),
\textsf{WEAT}, means are taken, which makes it insensitive to size
differences between \(X\) and \(Y\): \begin{align}
\mathsf{weat}(A,B) & = \frac{
\mu(\{\mathsf{s}(x,A,B)\}_{x\in X}) -\mu(\{\mathsf{s}(y,A,B)\}_{y\in Y}) 
}{
\sigma(\{\mathsf{s}(w,A,B)\}_{w\in X\cup Y})
}
\end{align}

\textsf{WEAT} is inspired by the Implicit Association Test (IAT)
{[}17{]} used in psychology, and it uses almost the same word sets,
allowing for a \emph{prima facie} sensible comparison with bias in
humans. In {[}8{]} the authors argue that significant biases---thus
measured--- similar to the ones discovered by IAT can be discovered in
word embeddings. In {[}11{]} the methodology is extended to a
multilingual and cross-lingual setting, arguing that using Euclidean
distance instead of cosine similarity does not make much difference,
while the bias effects vary greatly across embedding models
(interestingly, with social media-text trained embeddings being less
biased than those based on Wikipedia). A similar methodology is employed
in {[}4{]}. The authors employ word embeddings trained on corpora from
different decades to study the shifts in various biases.\footnote{Strictly
  speaking, these authors use Euclidean distances and their differences,
  but the way they take averages and averages thereof is analogous, and
  so what we will have to say about pre-averaging leading to false
  confidence applies to this methodology as well.}

\begin{figure}[H]

\begin{center}\includegraphics[width=0.55\linewidth]{paperDraft_files/figure-latex/WEATexample5-1} \end{center}
\caption{A simple example of a two-class set-up. Two groups of protected words, $X$ and $Y$ with two stereotypical attribute sets. An example of \textsf{WEAT} calculations follows.}
\label{eq:WEATexample}
\end{figure}

\noindent Here is an example of \textsf{WEAT} calculations for Figure
\ref{eq:WEATexample}:
\begin{align*} s_1 & = s(he,A,B)  =  \nicefrac{(.6+.7)}{2}  - \nicefrac{(.2+.1)}{2} = .65-.15= .5 \\
s_2  & = s(man,A,B) = .3 \\
s_3  & = s(woman,A,B) = -.6\\
s_4 & = s(she, A, B) = -.3\\
\mathsf{WEAT}(A,B)  & = \frac{\nicefrac{(s_1+s_2)}{2} - \nicefrac{(s_3+s_4)}{2}}{sd(\{s_1,s_2,s_3,s_4\})} \approx 1.93
\end{align*}

\textsf{WEAT}, however, has been developed to investigate biases
corresponding to a pair of supposedly opposing stereotypes, and so the
question arises as to how to generalize the measure to contexts in which
biases with respect to more than two stereotypical groups are to be
measured. Such a generalization can be found in {[}12{]}. The authors
introduce Mean Average Cosine distance (\textsf{MAC}) as a measure of
bias. Let \(T = \{t_1, \dots, t_k\}\) be a set of protected words, and
let each \(A_j\in \mathcal{A}\) be a set of attributes stereotypically
associated with a protected word where \(\mathcal{A}\). For instance,
when biases related to religion are to be investigated, they use a
dataset of the format illustrated in Table \ref{tab:religionOriginal}.
The measure is defined as follows:

\begin{table}
\footnotesize

\centering

\begin{tabular}[t]{lllr}
\toprule
protected words ($T$) & attributes & attribute set ($A_j$) & cosine distance\\
\midrule
\cellcolor{gray!15}{rabbi} & \cellcolor{gray!15}{greedy} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{1.03}\\
church & familial & christianStereotype & 0.70\\
\cellcolor{gray!15}{synagogue} & \cellcolor{gray!15}{liberal} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{0.79}\\
jew & familial & christianStereotype & 0.98\\
\cellcolor{gray!15}{quran} & \cellcolor{gray!15}{dirty} & \cellcolor{gray!15}{muslimStereotype} & \cellcolor{gray!15}{1.12}\\
muslim & uneducated & muslimStereotype & 0.52\\
\cellcolor{gray!15}{torah} & \cellcolor{gray!15}{terrorist} & \cellcolor{gray!15}{muslimStereotype} & \cellcolor{gray!15}{0.93}\\
quran & hairy & jewStereotype & 1.18\\
\cellcolor{gray!15}{synagogue} & \cellcolor{gray!15}{violent} & \cellcolor{gray!15}{muslimStereotype} & \cellcolor{gray!15}{0.95}\\
bible & cheap & jewStereotype & 1.22\\
\cellcolor{gray!15}{christianity} & \cellcolor{gray!15}{greedy} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{0.97}\\
muslim & hairy & jewStereotype & 0.88\\
\cellcolor{gray!15}{islam} & \cellcolor{gray!15}{critical} & \cellcolor{gray!15}{christianStereotype} & \cellcolor{gray!15}{0.79}\\
muslim & conservative & christianStereotype & 0.45\\
\cellcolor{gray!15}{mosque} & \cellcolor{gray!15}{greedy} & \cellcolor{gray!15}{jewStereotype} & \cellcolor{gray!15}{1.15}\\
\bottomrule
\end{tabular}

\caption{Sample 15 rows of the religion dataset. The whole dataset has 15 unique protected words ($T$), and 11 unique attributes divided between 3 attribute sets ($A_1=\mathsf{jewStereotype}, A_2= \mathsf{christianStereotype}, A_3=\mathsf{muslimStereotype}$). $\mathcal{A}$ consists of these three sets, $\mathcal{A}= \{A_1, A_2, A_3\}$. The whole dataset has $15\times 11 = 165$ rows.}
\label{tab:religionOriginal}
\normalsize 
\end{table}

\begin{align*}
\mathsf{s}(t, A_j) & = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cosineDistance}(t,a) \\
\mathsf{mac}(T,\mathcal{A}) & = \frac{1}{\vert T \vert \,\vert \mathcal{A}\vert}\sum_{t \in T }\sum_{A_j \in \mathcal{A}}    \mathsf{s}(t,A_j)
\end{align*}

\noindent That is, for each protected word \(t\in T\), and each
attribute set \(A_j\), they first take the mean of distances for this
protected word and all attributes in a given attribute class, and then
take the mean of thus obtained means for all the protected words and all
the protected classes.\footnote{The authors' code are in their github
  repository,
  \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding}.}

\begin{figure}[H]

\begin{center}\includegraphics[width=0.5\linewidth]{paperDraft_files/figure-latex/MACexample-1} \end{center}
\caption{A small subset of the religion dataset.  To each protected word in $T$  there corresponds one class of stereotypical attributes typically associated with it (and other classes of stereotypical attributes associated with different protected words).}
\label{fig:MACexample}
\end{figure}

\noindent An example of \textsf{MAC} calculations for the situation
depicted in Figure \ref{fig:MACexample} is as follows: \begin{align*}
s_1   = s(muslim,A_1)  & = \frac{\mathsf{cos}(muslim,dirty)+\mathsf{cos}(muslim,terrorist)}{2} \\ 
s_2   = s(muslim,A_2)  & = \frac{\mathsf{cos}(muslim,familiar)+\mathsf{cos}(muslim,conservative)}{2}\\ & \vdots \\ 
\mathsf{MAC}(T,A)  & = \mathsf{mean}(\{s_i \vert i \in 1, \dots, k\})
\end{align*}

Notably, the inutitive distinction between different attribute sets
plays no real role in the \textsf{MAC} calculations. Equally well one
could calculate the mean distance of \emph{muslim} to all the
predicates, mean distance of \emph{christian} to all the predicates,
mean distance of \emph{jew} to all the predicates, and then to take the
mean of these three means.

Having introduced the measures, first, we will introduce a selection of
general problems with this approach, and then we will move on to more
specific but important problems related to the fact that the measures
take averages and averages of averages. Once this is done, we move to
the development of our Bayesian alternative and the presentation of its
deployment.

\hypertarget{challenges-to-cosine-based-bias-metrics}{%
\section{Challenges to cosine-based bias
metrics}\label{challenges-to-cosine-based-bias-metrics}}

\label{sec:challenges}

\hypertarget{interpretability-issues}{%
\subsection{Interpretability issues}\label{interpretability-issues}}

\label{subsec:interpretability}

Here is an example of results of debiasing that can be found in {[}12{]}
(Table \ref{tab:religionOriginal2}). It presents the results of applying
the procedure with respect to the Religion dataset. \textsf{MAC} scores
are given for the original datasets, and two resulting from two
debiasing methods applied to the Reddit embedding. The scores for
debiasing come with p-values for the comparison to the original
\textsf{MAC}.

\begin{table}[H]
\footnotesize

\centering

\begin{tabular}[t]{lllr}
\toprule
Religion Debiasing & \textsf{MAC} & a p-value \\
\midrule
\cellcolor{gray!15}{Biased} & \cellcolor{gray!15}{0.859} & \cellcolor{gray!15}{N/A} \\
Hard Debiased & 0.934 & 3.006e-07\\
\cellcolor{gray!15}{Soft Debiased ($\lambda$ = 0.2)} & \cellcolor{gray!15}{0.894} & \cellcolor{gray!15}{0.007} \\
\bottomrule
\end{tabular}

\caption{The associated mean average cosine similarity
(\textsf{MAC}) and p-values for debiasing methods for religious bias.}
\label{tab:religionOriginal2}
\normalsize 
\end{table}

The first conceptual question we should ask is whether the initial
\textsf{MAC} values lower than 1 indeed are indicative of the presence
of bias? Of course, thinking abstractly, 1 is the ideal distance for
unrelated words. But clearly, there is variation in distances, which
might lead to non-biased lists also having \textsf{MAC} smaller than 1.
How much smaller? What may attract attention is the fact that the value
of cosine distance in ``Biased'' category is already quite high
(i.e.~close to 1) even before debiasing. High cosine distance indicates
low cosine similarity between values. One could think that average
cosine similarity equal to approximately 0.141 is not large enough to
consider it as bias. The authors, though, still aim to mitigate such
``bias'' to make the distance even larger. The question is, on what
basis is this small similarity still considered as a proof of the
presence of bias, and whether these small changes are meaningful.

The problem is that the original paper did not employ any control group
of neutral attributes for comparison to obtain a more realistic gauge on
how to understand \textsf{MAC} values. Later on, in our approach, we
introduce such control word lists. One of them is a list of words we
intuitively considered neutral (see the Appendix for word lists).

Moreover, it might be the case that words that have to do with human
activities in general, even if unbiased, are systematically closer to
the protected words than merely neutral words. This, again, casts doubt
on whether comparing \textsf{MAC} to the abstractly ideal value of 1 is
a methodologically sound idea. For this reason we also use a second list
with intuitively non-stereotypical human attributes (again, see the
Appendix for a word list).

Another important observation is that \textsf{MAC} calculations do not
distinguish whether a given attribute is associated with a given
protected word, simply averaging across all such groups. Let us use the
case of religion-related stereotypes to illustrate. The full lists from
{[}12{]} can be found in Appendix \ref{appendix:manzini_word_lists}. In
the original paper, words from all three religions were compared against
all of the stereotypes. No distinction between cases in which the
stereotype is associated with a given religion, as opposed to the
situation in which it is associated with another one, is made. For
example, the protected word \emph{jew} is supposed to be stereotypically
connected with the attribute \emph{greedy}, while from the protected
word \emph{quran} the attribute \emph{greedy} comes from a different
stereotype, and yet the distances between these pairs contribute equally
to the final \textsf{MAC} score. This is problematic, as not all of the
stereotypical words have to be considered as harmful for all of the
religions. To avoid the masking effect, one should pay attention to how
protected words and attributes are paired by stereotypes.

In Figures (\ref{fig:empirical0}-\ref{fig:empirical3}) we look at the
empirical distributions, while paying attention to such divisions. The
horizontal lines represent the values of 1-\textsf{MAC} (that is, we now
talk in terms of cosine similarity rather than cosine distance; this is
a trivial linear transformation) that the authors considered indicative
of bias for stereotypes corresponding to given word lists. For instance,
in religion, \textsf{MAC} was .859 and considered a sign of bias, so we
plot \(0\pm (1-.859)\approx .14\) lines around similarity = 0 (that is,
distance = 1). Notice that most distributions are quite wide, and the
proportions of even neutral or human neutral words with similarities
higher than the value of 1-\textsf{MAC} deserving debiasing according to
the authors is quite high.

\begin{figure}[H]

\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8d-1} \end{center}

\caption{Empirical distributions of cosine similarities  for the Religion word list  used in  the original paper. }

\label{fig:empirical0}
\end{figure}

\begin{figure}[H]

\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8df-1} \end{center}

\caption{Empirical distributions of cosine similarities for the Gender word list used in  the original paper.  }

\label{fig:empirical2}
\end{figure}

\begin{figure}[H]

\begin{center}\includegraphics[width=0.7\linewidth]{paperDraft_files/figure-latex/fig:cosineDistributions8dg-1} \end{center}

\caption{Empirical distributions of cosine similarities  for the Race word list used in  the original paper.  }

\label{fig:empirical3}
\end{figure}

Another issue to consider is the selection of attributes for bias
measurement. The word lists used in the literature are often fairly
small (5-50). The papers in the field do employ statistical tests to
measure the uncertainty involved and do make claims of statistical
significance. Yet, we will later on argue that these method are not
proper for the goal at hand. By using Bayesian methods we will show that
a more appropriate use of statistical methods leads to estimates of
uncertainty which suggest that larger word lists would be advisable.

To avoid such problems, later on, we employ control groups and in line
with Bayesian methodology, use posterior distributions and highest
posterior density intervals instead of chasing single-point metrics
based on pre-averaged data. But before we do so, let us spend some time
explaining why pre-averaging is a sub-optimal strategy.

\hypertarget{problems-with-pre-averaging}{%
\subsection{Problems with
pre-averaging}\label{problems-with-pre-averaging}}

\label{subsec:problems}

The approaches we have been describing use means of mean average cosine
similarities to measure similarity between protected words and
attributes coming from harmful stereotypes. However, a closer inspection
look at the individual values that are taken for the calculations, it
turns out that there are quite a few outliers and surprisingly
dissimilar words. This problem becomes transparent when we examine the
visualizations of the individual cosine distances, following the idea
that one of the first steps in understanding data is to look at it.
Let's start with inspecting two examples in Figures \ref{fig:muslim} and
\ref{fig:priest}.

\begin{figure}[H]

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visMuslimd-1} \end{center}
\caption{Actual distances for the protected word "muslim". The most similar attributes tend to be the ones associated with it stereotypically, but then words associated with other stereotypes come closer than neutral or human predicates. }
\label{fig:muslim}
\end{figure}

\begin{figure}[H]

\begin{center}\includegraphics[width=1\linewidth]{paperDraft_files/figure-latex/visPriestd-1} \end{center}

\caption{Actual distances for the protected word "priest", the situation is even less as expected: the nearest attributes are human attributes, and all there seems to be no clear pattern to the distances to other attributes.}
\label{fig:priest}
\end{figure}

With such a method the uncertainty involved is not really considered
which makes it even more difficult to give reasonable interpretations of
the results. We propose the use of Bayesian method to obtain some
understanding of the influence the uncertainty has on the interpretation
of final results.

\noindent \(s(X,Y,A,B)\) is the statistic used in the significance test,
and the \(p\)-value is obtained by bootstrapping: it is the frequency of
\(s(X_i,Y_i,A,B)>s(X,Y,A,B)\) for all equally sized partitions
\(X_i, Y_i\) of \(X\cup Y\). The effect size is computed by normalizing
the difference in means as follows:

\vspace{-2mm}

\footnotesize

\begin{align}
bias(A,B) & = \frac{
\mu(\{s(x,A,B)\}_{x\in X}) -\mu(\{s(y,A,B)\}_{y\in Y}) 
}{
\sigma(\{s(w,A,B)\}_{w\in X\cup Y})
}
\end{align}

\normalsize

The t-tests they employ are run on average cosines used to calculate
\textsf{MAC}.

The effect sizes reported by {[}8{]} for lists of words for which the
embeddings are supposedly biased range from 2.06 to 1.81, and the
reported \(p\)-values are in the range of \(10^{-7}-10^{-2}\) with one
exception for Math vs arts, where it is \(.018\). The question is, are
those results meaningful?

One way to answer this question is to think in terms of null models. If
the words actually are samples from two populations with equal mean, how
often would we see effect sizes in this range? How often would we reach
the \(p\)-values that the authors reported?

We start our exploration with considering a list of 16 protected words,
each class containing 8, with 16 attributes, 8 in each attribute class.
In each particular sample we will investigate the consequence of the
model being in fact null: we draw the cosine distances, every time, from
the \(\mathsf{Normal}(0,.08)\) distribution. \(0.08\) is approximately
the empirical standard deviation observed in fairly large samples of
neutral words, and 16 is the sample size used in the WEAT7 word list,
which is not much different from the other sample sizes in word lists
used by {[}8{]}

So let's draw one sample of this type. We calculate the s-values, the
test statistics and the effect sizes. Then, following the original
methodology, we obtain bootstrapped distributions of the effect sizes
and the test statistics by calculating the same for each possible equal
split of the initial random data set. The observed test statistic is
0.386222 and 1.2671948. The bootstrapped distributions of the test
statistics and effect sizes are illustrated in Figure
\ref{fig:caliskanCalc}. Quite notably both (two-sided) \(p\) values are
uneven and rather low.

\begin{figure}[H]

\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:caliskanCalcB2B-1} \end{center}

\caption{Bootstrapped distributions of test statistics and effect sizes in a random sample given the null hypothesis.}
\label{fig:caliskanCalc}
\end{figure}

At this point, we might think---right, so while drawing on the
assumption of the null hypothesis we just stumbled into a data set that
randomly happened to display strong bias. We decide to double-check this
by visual inspection expecting exactly this: a strong, clearly visible
bias (Figure \ref{fig:caliskanDistances}).

\begin{figure}[H]

\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:caliskanDistC2B-1} \end{center}

\caption{Cosine distances to two attribute sets by protected word groups. Observe nothing unusual except for a few outliers.}
\label{fig:caliskanDistances}
\end{figure}

\noindent In fact, while there might be some outliers here and there,
saying that a clear bias on which one group is systematically closer to
\(A\)s than another is definitely a stretch. What happened?

In the calculations of \textsf{WEAT} means are taken twice. The
\(s\)-values themselves are means, and then means of \(s\)-values are
compared between groups. The test statistic used itself is of this sort
as well---while it compares sums, these are sums of sets of the same
size, so they are really linear transformations of means of means.
Statistical troubles start when we run statistical tests on sets of
means, for at least two reasons One, by pre-averaging data we throw away
information about sample sizes. For the former point, think about
proportions: 10 out of 20 and 2 out of 4 give the same mean, but you
would obtain more information by making the former observation rather
than by making the latter. And especially in this context, in which the
word lists are not huge, sample sizes should matter. Two, when we
pre-average, we remove variation, and so pre-averaging tend to
manufacture false confidence. Group means display less variation than
the raw data points, and the standard deviation of a set of means of
sets of means is bound to be lower than the original standard deviation
in the row data. Now, if you calculate your effect size by dividing by
the pre-averaged standard deviation, your quite likely to get something
that looks like a strong effect size, but the results of your
calculations might not track anything interesting.

So let us think again about the question that we are ultimately
interested in. Are the \(X\) terms systematically closer to (further
from) the \(A\) attributes (\(B\) attributes) than the \(Y\) words? But
now let's use the raw data points to try to answer these questions. As
the first stab, let us run two quick \(t\)-tests to gauge what the raw
data illustrated in Figure \ref{fig:caliskanDistances} tell us.

First, distances to \(A\) attributes for \(X\) words and \(Y\) words.
Whoa, the result is significant! The \(p\)-value is \(0.02\). So sure,
the sample is in some sense unusual. But the 95\% confidence interval
for the difference in means is \([.0052 .061]\), clearly nothing that a
reader would expect given that the calculated effect size seemed quite
large. How about the \(B\) attributes? Here the \(p\)-value is \(.22\)
and the 95\% confidence interval is \([-0.03, .009]\), even less of a
reason to think a bias is present.

The difficulties are exacerbated by the fact that statistical tests are
based on bootstrapping from a relatively small data sets, which is quite
likely to underestimate the population variance. To make our point
clear, let us avoid bootstrapping and work with the null generative
model with \(\mathsf{Norm}(0,.08)\) for both word groups. We keep the
sizes the same: we have eight protected words in each group, sixteen in
total, and for each we randomly draw 8 distances from hypothetical \(A\)
attributes, and \(8\) distances from hypothetical \(B\) attributes.
Calculate the test statistic and effect size the way {[}8{]} did. Do
this 10000 times and look at what the distributions of these values are
on the assumption of the null model with realistic empirically motivated
raw data point standard deviation.

\begin{figure}[H]

\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/fig:ourDistancesPlot2e-1} \end{center}

\caption{Distributions of test statistics and effect sizes based on 10k simulations on the assumption of a null model in which all distances come from normal distribution with $\mu =0, \sigma = .08, n=10k$.}
\label{fig:ourDistances}
\end{figure}

The first observation is that the supposedly large effect size we
obtained is not that unusual even assuming a null model. Around 38\% of
samples result in effect size at least as extreme. This illustrates the
point that the effect size used does not track anything interesting.
Second, the distribution of test statistics is much more narrow, which
means that if we use it to calculate \(p\)-values, it is not too
difficult to obtain a supposedly significant test statistic which
nevertheless does not correspond to anything interesting happening in
the data set.

Where do we stand now? We have seen that seemingly high effect sizes
might arise even if the underlying processes actually have the same
mean. The uncertainty resulting from including the raw data point
variance into considerations is more extensive than the one suggested by
the low p-values obtained from taking means or means of means as data
points. In this section, we discussed the performance of the
\textsf{WEAT} measure, but since the {[}12{]} one is a generalization
thereof, including the method of running statistical tests on
pre-averaged data, our remarks, \emph{mutatis nutandis}, apply.
Moreover, a general problem is that studies such as {[}8{]} or {[}12{]}
do not use control groups, and this makes the interpretation of the
analysis additionally difficult.

What is the alternative? As we already emphasized: focusing on what the
real underlying question is, and trying to answer it using a statistical
analysis of the raw data. Moreover, since the data sets are not too
large and since multiple evaluations are to be made, we will pursue this
method from the Bayesian perspective.

\hypertarget{bayesian-method}{%
\section{Bayesian method}\label{bayesian-method}}

\label{sec:bayesian}

\hypertarget{model-construction}{%
\subsection{Model construction}\label{model-construction}}

\label{subsec:model}

Bayesian data analysis takes prior probability distributions, a
mathematical model structure and the data, and returns the posterior
probability distributions over the parameters of interest, thus
capturing our uncertainty about their actual values. One important
difference between such a result and the result of a classical
statistical analysis is that classical confidence intervals (CIs) have a
rather complicated and somewhat confusing interpretation, which has
little to do with the posterior probability distribution.\footnote{Here
  are a few usual problems. CIs are often mistakenly interpreted as
  providing the probability that a resulting confidence interval
  contains the true value of a parameter. CIs bring confusion also with
  regard to precision, it is a common mistake to interpret narrow
  intervals as the ones corresponding to a more precise knowledge.
  Another fallacy is to associate CIs with likelihood and stating that
  values within a given interval are more probable than the ones outside
  it. The theory of confidence intervals does not support the above
  interpretations. CIs should be plainly interpreted as a result of
  certain procedure (there are many ways to obtain CIs from a given set
  of data) that will in the long run contain the true value if the
  procedure is performed a fixed amount of times. For a nice survey and
  explanation of these misinterpretations, see {[}15{]}. For a
  psychological study of the occurrence of such misinterpretations, see
  REF. In this study, 120 researchers and 442 students were asked to
  assess the truth value of six false statements involving different
  interpretations of a CI. Both researchers and students endorsed, on
  average, more than three of these statements.} {[}7{]}

In fact, Bayesian highest posterior density intervals (HPDIs, the
narrowest intervals containing a certain ratio of the area under the
curve) and CIs end up being numerically the same only if the prior
probabilities are uniform. This illustrates that (1) classical analysis
is unable to incorporate non-trivial priors, and (2) is therefore more
susceptible to over-fitting, unless regularization (equivalent to a more
straightforward Bayesian approach) is used. In contrast with CIs, the
posterior distributions are easily interpretable and have direct
relevance for the question at hand. Moreover, Bayesian data analysis is
better at handling hierarchical models and small datasets, which is
exactly what we will we dealing with.

In a standard Bayesian analysis, the first step is to understand the
data and select potential predictors and the outcome variable. Once the
data is understood, the next step is to formulate a mathematical
description of the generative model of the relationships between the
predictors and the outcome variable. Prior distributions must then be
chosen for the parameters used in the model. Next, Bayesian inference
must be applied to find posterior probabilities over the possible
parameter values. Finally, after finding the posterior probabilities, it
is important to check how well the posterior predictions reflect the
data with a posterior predictive check. In our analysis the outcome
variable is the cosine distances between the protected words and
attribute words. The predictor is a factor determining whether a given
attribute word is a neutral word, a human predicate, is stereotypically
associated with the protected word, or comes from a different stereotype
connected with another protected word. The idea is, if bias is present
in the embedding, distances to associated attribute words should be
systematically lower than to other attribute words.

Furthermore, conceptually there are two levels of analysis in our
approach (see Figure \ref{fig:visDag}). On the one hand, we are
interested in the general question of whether related attributes are
systematically closer across the dataset. On the other hand, we are
interested in a more fine-grained picture of the role of the predictor
for particular protected words. Learning in hierarchical Bayesian models
involves using Bayesian inference to update the parameters of the model.
This update is based on the observed data, and estimates are made at
different levels of the data hierarchy. We use hierarchical Bayesian
models in which we simultaneously estimate parameters at the protected
word level and at the global level, assuming that all lower-level
parameters are drawn from global distributions. Such models can be
thought of as incorporating adaptive regularization, which avoids
overfitting and leads to improved estimates for unbalanced datasets (and
the datasets we need to use are unbalanced).

\begin{figure}[H]

\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/dagVis-1} \end{center}

\caption{At a general level, we will be estimating the coefficients for distances as grouped by whether they are between protected words and attributes coming from their respective associated/different/human/neutral attribute groups. At a more fine-grained level, for each protected word we will be estimating the proximity of that word to attributes that are associated with its respective stereotype, come from a different stereotype, or come from the human/neutral attribute lists.}
\label{fig:visDag}
\end{figure}

To be more specific, the underlying mathematical model is as follows.
First, we assume that distances are normally distributed:
\begin{align*} \mathsf{distance_i} & \sim \mathsf{dnorm}(\mu_i,\sigma_i)
\end{align*} \noindent Second, for each particular protected word
\(\mathsf{pw}\) there are four parameters to be estimated. Its mean
distance to associated stereotypes \(a[pw]\), its mean distance to
attributes coming from different stereotypes, \(d[pw]\), its mean
distance to human attributes, \(h[pw]\), and its mean distance to
neutral attributes, \(n[pw]\): \begin{align*}
\mu_i & = d_{\mathsf{pw[i]}} \times \mathsf{different}_i  + a_{\mathsf{pw[i]}} \times \mathsf{associated}_i  + h_{\mathsf{pw[i]}} \times \mathsf{human}_i  + n_{\mathsf{pw[i]}}\times \mathsf{neutral}_i
\end{align*} \noindent where
\(\mathsf{different}, \mathsf{associated},\mathsf{human}\) and
\(\mathsf{neutral}\) are binary variables.

This completes our description of the underlying mechanism. Now the
priors and the hierarchy. We assume all the \(a\) parameters come from
one distribution, that is normal around a higher-level parameter
\(\bar{a}\) and so on for the other three groups of parameters. That is,
\(a_{\mathsf{pw[i]}}\) is the average distance of a given particular
protected word to attributes stereotypically associated with it, while
\(\bar{a}\) is the overall average distance of protected words to
attributes associated with them {[}10,13{]}. \begin{align*}
d_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{d}, \overline{\sigma_d}) \\
a_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{a}, \overline{\sigma_a}) \\
h_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{h}, \overline{\sigma_h}) \\
n_{\mathsf{pw[i]}} &\sim \mathsf{Norm}(\bar{n}, \overline{\sigma_n}) 
\end{align*}

According to our priors, the group means \(\bar{a}\), \(\bar{d}\),
\(\bar{h}\) and \(\bar{n}\) all come from one normal distribution with
mean equal to \(1\) and standard deviation equal to \(.3\). The standard
deviations \(\bar{\sigma_a}\), \(\bar{\sigma_d}\), \(\bar{\sigma_h}\)
and \(\bar{sigma_n}\) to be estimated, according to our prior, come from
one distribution, exponential with rate parameter equal to \(2\). Our
priors are slightly skeptical. They do reflect our knowledge and
intuition on the probable distribution of the cosine distances in the
data. We know that the cosine distances lie in the range \(0-2\), and we
expect two randomly chosen vectors from the embedding to have rather
small similarity, so we expect the distances to be centered around
\(1\). However, we use a rather wide standard deviation (\(.3\)) to
easily account for cases where there is actually much higher similarity
between two vectors (especially in cases where the embedding is supposed
to be biased). Our priors for the standard deviations are also fairly
weak.\footnote{To build a model one may use original \textit{Stan} which is a state-of-the-art platform for statistical modeling and high-performance statistical computation (https://mc-stan.org/). The simpler option is to use \textit{Rethinking} package from the \textit{Statistical rethinking} course. With one of the functions \textit{ulam} one may build RStan models from the formulas in an easy manner}

\begin{align*}
\bar{d}, \bar{a}, \bar{h}, \bar{n} &\sim \mathsf{Norm}(1, .3)\\ 
\overline{\sigma_d}, \overline{\sigma_a},  \overline{\sigma_h},  \overline{\sigma_n}  &\sim \mathsf{Exp}(2)
\end{align*}

\hypertarget{posterior-predictive-check}{%
\subsection{Posterior predictive
check}\label{posterior-predictive-check}}

\label{subsec:posterior}

A posterior predictive check is a technique used to evaluate the fit of
a Bayesian model by comparing its predictions with observed data. The
underlying principle is to generate simulated data from the posterior
distribution of the model parameters and compare them with the observed
data. If the model is a good fit to the data, the simulated data should
resemble the observed data. A posterior predictive check is crucial
because it provides a way to assess the validity of a Bayesian model.
This is important because a model that fits the observed data well may
still fail to generalize to new data. In the figure below we can see
that the model fits the the data well.

\begin{figure}[H]

\begin{center}\includegraphics[width=0.8\linewidth]{paperDraft_files/figure-latex/figposteriorPrCheckA6b-1} \end{center}
\caption{Example of a posterior predictive check. (Left) Actual cosine distances are plotted against mean predictions with 89\% highest posterior density intervals. Notice that 90\% of actual values fall within the 89\% HPDI and 55\% of actual values fall into 50\% HPDI, which indicates appropriate performance of the model. The left-right alignment of differrent colors coressponds to the fact that cosine differences between elements of different categories differ, to some extent systematically (this will be studied in the results section). (Right) Densities of predicted and observed distances.}
\label{fig:posteriorCheck1}
\end{figure}

\normalsize

\hypertarget{results-and-discussion}{%
\section{Results and discussion}\label{results-and-discussion}}

\label{sec:results}

\vspace{1mm}

\normalsize

\hypertarget{observations}{%
\subsection{Observations}\label{observations}}

\label{subsec:observations}

In brief, despite one-number metrics suggesting otherwise, our Bayesian
analysis reveals that insofar as the short word lists usually used in
related research projects are involved, there usually are no strong
reasons to claim the presence of systematic bias. Moreover, comparison
between the groups (including control word lists) leads to the
conclusion that the effect sizes (that is, the absolute differences
between cosine distances between groups) tend to be rather small, with
few exceptions. Moreover, the choice of protected words is crucial ---
as there is a lot of variance when it comes to the protected word-level
analysis.

In a bit more detail, the visualizations in Appendix
\ref{appendix:visualizations} show that the situation is more
complicated than merely looking at one-number summaries might suggest.
Note that the axes are sometimes in different scales to increase
visibility.

To start with, let us look at the association-type level coefficients
(illustrated in the top parts of the plots). Depending on the corpus
used and word class, there is a large variety as to the posterior
densities. Quite aware of this being a crude approximation, let's
compare the HPDIs and whether they overlap for different attribute
groups.

\begin{itemize}
\item
  In Weat 7 (Reddit) there is no reason to think there are systematic
  differences between cosine distances (recall that words from Weat 7
  were mostly not available in other embeddings).
\item
  In Weat 1 (Google, Glove and Reddit) associated words are somewhat
  closer, but the cosine distance differences from neutral words are
  very low, and surprisingly it is human attributes, not neutral
  predicates that are systematically the furthest.
\item
  In Religion (Google, Glove, Reddit) and Race (Google, Glove), the
  associated attributes are not systematically closer than attributes
  belonging to different stereotypes, and the difference from neutral
  and human predicates is rather low, if noticeable. The situation is
  interestingly different in Race (Reddit) where both human and neutral
  predicates are systematically further than associated and different
  attributes - but even then, there is no clear difference between
  associated and different attributes.
\item
  For Gender (Google, Glove), despite the superficially binary nature,
  associated and opposite attributes tend to be more or less in the same
  distances, much closer than neutral words (but not closer than human
  predicates in Glove). Reddit is an extreme example: both associated
  and opposite attributes are much closer than neutral and human (around
  .6 vs.~.9), but even then, there seems to be no reason to think than
  cosine distances to associated predicates are much different from
  distances to opposite predicates.
\end{itemize}

Moreover, when we look at particular protected words, the situation is
even less straightforward. We will just go over a few notable examples,
leaving the visual inspection of particular results for other protected
words to the reader. One general phenomenon is that--as we already
pointed out--the word lists are quite short, which contributes to large
uncertainty involved in some cases.

\begin{itemize}
\item
  For some protected words the different attributes are somewhat closer
  than the associated attributes.
\item
  For some protected words, associated and different attributes are
  closer than neutral attributes, but so are human attributes.
\item
  In some cases, associated attributes are closer, but so are neutral
  and human predicates, which illustrates that just looking at average
  cosine similarity as compared to the theoretically expected value of
  1, instead of running comparison to neutral and human attributes is
  misleading.
\item
  The only group of protected words where differences are noticeable at
  the protected word level is Gender-related words-- as in Gender
  (Google) and in Gender (Reddit) --- note however that in the latter,
  for some words the opposite attributes seem to be a little bit closer
  than the associated ones.
\end{itemize}

\hypertarget{rethinking-debiasing}{%
\subsection{Rethinking debiasing}\label{rethinking-debiasing}}

\label{subsec:rethinking}

These visualizations can be also handy when it comes to the
investigation of the effect that debiasing has on the embedding space.
In Figures \ref{fig:empiricalPriorToDebiasing} and
\ref{fig:empiricalDebiased} we see chosen pair depicting the difference
in means with 89\% highest posterior density intervals before and after
applying debiasing.

\begin{figure}[H]

\begin{center}\includegraphics{paperDraft_files/figure-latex/fig:debiasedCosinePair2e-1} \end{center}
\caption{Mean cosine distances with 89\% highest posterior density intervals for the  gender dataset before debiasing.}
\label{fig:empiricalPriorToDebiasing}
\end{figure}

\begin{figure}[H]

\begin{center}\includegraphics{paperDraft_files/figure-latex/fig:debiasedCosinePair2_e-1} \end{center}
\caption{Mean with 89\% highest posterior density intervals for gender after debiasing.}
\label{fig:empiricalDebiased}
\end{figure}

In Figures \ref{fig:empiricalDebiased}-\ref{fig:empiricalDebiased3} we
inspect the empirical distributions for the debiased embeddings.
Comparing the results to the original embedding, one may notice that for
the Religion group the neutral and human distribution has changed
slightly. Before within the ``correct'' cosine similarity boundaries
there were 56\% of neutral and 55\% of human word lists. After the
debiasing the values changed to 59\% (for neutral) and 59\% (for human).
The different and associated word lists were more influenced. The
general shape of both distributions is less stretched. Before debiasing
43\% of the different word lists and 35\% of the associated word lists
were within the accepted boundaries. After the embedding manipulation
the percentage has increased for both lists to 63\%. Visualization for
Gender group illustrates almost no change for the neutral and human word
lists before and after debiasing. The values for different and
associated word lists are also barely impacted by the embedding
modification. In the Race group, the percentage within the boundaries
for neutral and associated word lists have increased. The opposite
happened for human and different word lists, where the percentage of
``correct'' cosine similarity dropped from 67\% to 55\% (human) and from
39\% to 36\% (different).

\begin{figure}[H]

\begin{center}\includegraphics[width=0.95\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine21c-1} \end{center}
\caption{Empirical distributions of cosine similarities before and after  debiasing  for  the Religion word list  used in  the original paper.}
\label{fig:empiricalDebiased}
\end{figure}

\begin{figure}[H]

\begin{center}\includegraphics[width=0.95\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine22c-1} \end{center}
\caption{Empirical distributions of cosine similarities before and after  debiasing for  the Gender word list used in  the original paper.}
\label{fig:empiricalDebiased2}
\end{figure}

\begin{figure}[H]

\begin{center}\includegraphics[width=0.95\linewidth]{paperDraft_files/figure-latex/fig:debiasedCosine23c-1} \end{center}
\caption{Empirical distributions of cosine similarities before and after  debiasing for the Race word list used in  the original paper.}
\label{fig:empiricalDebiased3}
\end{figure}

\hypertarget{related-work-and-conclusions}{%
\section{Related work and
conclusions}\label{related-work-and-conclusions}}

\label{sec:related}

It has been suggested {[}2{]} to use Bernstein bounds to state the
uncertainty of the sample bias in a form of confidence interval. The
interpretation of this interval is as follows, 95\% confidence interval
contains such a range of values, that one can be 95\% certain that it
contains the population mean. As we will prove later in the text,
confidence intervals are quite problematic for several reasons, among
others the confusing interpretability. In {[}2{]} the current problem of
bias estimation was correctly identified. Indeed, the bias estimate
should not be expressed as a single number without taking into account
that the estimate is made on a sample of data and therefore has
intrinsic uncertainty. However, we argue that Bernstein bounds do not
provide the best solution to this problem. Applying their method to a
popular WinoBias dataset leads to the conclusion that more than 11903
samples are needed to claim a 95\% confidence interval for a bias
estimate. This amount exceeds existing datasets for bias measurement. We
propose a more realistic method that can be applied to commonly used
datasets to estimate the uncertainty of the bias measurement.

\todo{More references to similar papers}

A Bayesian data analysis with hierarchical models of cosine distances
between protected words, control group words, and stereotypical
attributes provides more modest and realistic assessment of the
uncertainty involved. It reveals that much complexity is hidden when one
instead chases single bias metrics present in the literature.

After introducing the method, we apply it to multiple word embeddings
and results of supposed debiasing, putting forward some general
observations that are not exactly in line with the usual picture painted
in terms of \textsf{WEAT} or \textsf{MAC} (and the problem generalizes
to any approach that focuses on chasing a single numeric metric): the
word list sizes and sample sizes used in the studies are usually small.
posterior density intervals are fairly wide, often the differences
between associated, different and neutral human predicates, are not very
impressive. Also, a preliminary inspection suggests that the
desirability of changes obtained by the usual debiasing methods is
debatable. The tools that we propose, however, allow for a more
fine-grained and multi-level evaluation of bias and debiasing in
language models without losing modesty about the uncertainties involved.

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Bolukbasi2016man}{}}%
\CSLLeftMargin{{[}1{]} }%
\CSLRightInline{Tolga Bolukbasi, Kai-Wei Chang, James Y. Zou, Venkatesh
Saligrama, and Adam Kalai. 2016. Man is to computer programmer as woman
is to homemaker? Debiasing word embeddings. \emph{CoRR} abs/1607.06520,
(2016). Retrieved from \url{http://arxiv.org/abs/1607.06520}}

\leavevmode\vadjust pre{\hypertarget{ref-Ethayarajh2020Bernstein}{}}%
\CSLLeftMargin{{[}2{]} }%
\CSLRightInline{Kawin Ethayarajh. 2020. Is your classifier actually
biased? Measuring fairness under uncertainty with bernstein bounds.
\emph{CoRR} abs/2004.12332, (2020). Retrieved from
\url{https://arxiv.org/abs/2004.12332}}

\leavevmode\vadjust pre{\hypertarget{ref-Garg2017hundredYears}{}}%
\CSLLeftMargin{{[}3{]} }%
\CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
Zou. 2017. Word embeddings quantify 100 years of gender and ethnic
stereotypes. \emph{Proceedings of the National Academy of Sciences} 115,
(November 2017).
DOI:https://doi.org/\href{https://doi.org/10.1073/pnas.1720347115}{10.1073/pnas.1720347115}}

\leavevmode\vadjust pre{\hypertarget{ref-Garg2018years}{}}%
\CSLLeftMargin{{[}4{]} }%
\CSLRightInline{Nikhil Garg, Londa Schiebinger, Dan Jurafsky, and James
Zou. 2018. Word embeddings quantify 100 years of gender and ethnic
stereotypes. \emph{Proceedings of the National Academy of Sciences} 115,
16 (April 2018), E3635--E3644.
DOI:https://doi.org/\href{https://doi.org/10.1073/pnas.1720347115}{10.1073/pnas.1720347115}}

\leavevmode\vadjust pre{\hypertarget{ref-Gonen2019lipstick}{}}%
\CSLLeftMargin{{[}5{]} }%
\CSLRightInline{Hila Gonen and Yoav Goldberg. 2019. Lipstick on a pig:
{D}ebiasing methods cover up systematic gender biases in word embeddings
but do not remove them. In \emph{Proceedings of the 2019 conference of
the north {A}merican chapter of the association for computational
linguistics: Human language technologies, volume 1 (long and short
papers)}, Association for Computational Linguistics, Minneapolis,
Minnesota, 609--614.
DOI:https://doi.org/\href{https://doi.org/10.18653/v1/N19-1061}{10.18653/v1/N19-1061}}

\leavevmode\vadjust pre{\hypertarget{ref-gordon2012reporting}{}}%
\CSLLeftMargin{{[}6{]} }%
\CSLRightInline{Jonathan Gordon and Benjamin Durme. 2013. Reporting bias
and knowledge acquisition. In \emph{AKBC 2013 - Proceedings of the 2013
Workshop on Automated Knowledge Base Construction, Co-located with CIKM
2013}, 25--30.
DOI:https://doi.org/\href{https://doi.org/10.1145/2509558.2509563}{10.1145/2509558.2509563}}

\leavevmode\vadjust pre{\hypertarget{ref-Hoekstra2014Misinterpretation}{}}%
\CSLLeftMargin{{[}7{]} }%
\CSLRightInline{Rink Hoekstra, Richard D. Morey, Jeffrey N. Rouder, and
Eric-Jan Wagenmakers. 2014. Robust misinterpretation of confidence
intervals. \emph{Psychonomic Bulletin \& Review} 21, 5 (October 2014),
1157--1164.
DOI:https://doi.org/\href{https://doi.org/10.3758/s13423-013-0572-3}{10.3758/s13423-013-0572-3}}

\leavevmode\vadjust pre{\hypertarget{ref-Caliskan2017semanticsBiases}{}}%
\CSLLeftMargin{{[}8{]} }%
\CSLRightInline{Aylin Caliskan Islam, Joanna J. Bryson, and Arvind
Narayanan. 2016. Semantics derived automatically from language corpora
necessarily contain human biases. \emph{CoRR} abs/1608.07187, (2016).
Retrieved from \url{http://arxiv.org/abs/1608.07187}}

\leavevmode\vadjust pre{\hypertarget{ref-JohnsonValueFree}{}}%
\CSLLeftMargin{{[}9{]} }%
\CSLRightInline{Gabbrielle Johnson. forthcoming. Are algorithms
value-free? Feminist theoretical virtues in machine learning.
\emph{Journal Moral Philosophy} (forthcoming).}

\leavevmode\vadjust pre{\hypertarget{ref-kruschke2015bayesian}{}}%
\CSLLeftMargin{{[}10{]} }%
\CSLRightInline{John Kruschke. 2015. \emph{Doing bayesian data analysis
(second edition)}. Academic Press, Boston.}

\leavevmode\vadjust pre{\hypertarget{ref-Lauscher2019multidimensional}{}}%
\CSLLeftMargin{{[}11{]} }%
\CSLRightInline{Anne Lauscher and Goran Glavas. 2019. Are we
consistently biased? Multidimensional analysis of biases in
distributional word vectors. \emph{CoRR} abs/1904.11783, (2019).
Retrieved from \url{http://arxiv.org/abs/1904.11783}}

\leavevmode\vadjust pre{\hypertarget{ref-Manzini2019blackToCriminal}{}}%
\CSLLeftMargin{{[}12{]} }%
\CSLRightInline{Thomas Manzini, Yao Chong Lim, Yulia Tsvetkov, and Alan
W Black. 2019. Black is to criminal as caucasian is to police: Detecting
and removing multiclass bias in word embeddings. Retrieved from
\url{https://arxiv.org/abs/1904.04047}}

\leavevmode\vadjust pre{\hypertarget{ref-statrethinkingbook2020}{}}%
\CSLLeftMargin{{[}13{]} }%
\CSLRightInline{Richard McElreath. 2020. \emph{Statistical rethinking: A
bayesian course with examples in r and stan, 2nd edition} (2nd ed.). CRC
Press. Retrieved from
\url{http://xcelab.net/rm/statistical-rethinking/}}

\leavevmode\vadjust pre{\hypertarget{ref-Mikolov2013efficient}{}}%
\CSLLeftMargin{{[}14{]} }%
\CSLRightInline{Tomas Mikolov, Kai Chen, Greg Corrado, and Jeffrey Dean.
2013. Efficient estimation of word representations in vector space.
DOI:https://doi.org/\href{https://doi.org/10.48550/ARXIV.1301.3781}{10.48550/ARXIV.1301.3781}}

\leavevmode\vadjust pre{\hypertarget{ref-Morey2015confidenceFallacy}{}}%
\CSLLeftMargin{{[}15{]} }%
\CSLRightInline{Richard Morey, Rink Hoekstra, Jeffrey Rouder, Michael
Lee, and EJ Wagenmakers. 2015. The fallacy of placing confidence in
confidence intervals. \emph{Psychonomic Bulletin \& Review} (September
2015).}

\leavevmode\vadjust pre{\hypertarget{ref-Nissim2020fair}{}}%
\CSLLeftMargin{{[}16{]} }%
\CSLRightInline{Malvina Nissim, Rik van Noord, and Rob van der Goot.
2020. Fair is better than sensational: Man is to doctor as woman is to
doctor. \emph{Computational Linguistics} 46, 2 (June 2020), 487--497.
DOI:https://doi.org/\href{https://doi.org/10.1162/coli_a_00379}{10.1162/coli\_a\_00379}}

\leavevmode\vadjust pre{\hypertarget{ref-Nosek2002harvesting}{}}%
\CSLLeftMargin{{[}17{]} }%
\CSLRightInline{Brian A. Nosek, Mahzarin R. Banaji, and Anthony G.
Greenwald. 2002. Harvesting implicit group attitudes and beliefs from a
demonstration web site. \emph{Group Dynamics: Theory, Research, and
Practice} 6, 1 (2002), 101--115.
DOI:https://doi.org/\href{https://doi.org/10.1037/1089-2699.6.1.101}{10.1037/1089-2699.6.1.101}}

\end{CSLReferences}

\newpage
\appendix

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

\label{sec:appendix}

\hypertarget{visualizations}{%
\subsection{Visualizations}\label{visualizations}}

\label{appendix:visualizations}

\begin{center}\includegraphics{paperDraft_files/figure-latex/weatjoint2z-1} \end{center}

\pagebreak

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/weat7glovh-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat7Reddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Googlea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Glovea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsWeat1Reddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGooglea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionGlovea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsReligionReddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGooglea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceGlovea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsRaceReddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGooglea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderGlovea-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsGenderReddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedReligionReddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedRaceReddita-1} \end{center}

\begin{center}\includegraphics[width=1.1\linewidth]{paperDraft_files/figure-latex/resultsDebiasedGenderReddita-1} \end{center}

\hypertarget{posterior-predictive-checks}{%
\subsection{Posterior predictive
checks}\label{posterior-predictive-checks}}

\label{appendix:posterior}

\begin{center}\includegraphics[width=0.85\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixa-1} \end{center}

\begin{center}\includegraphics[width=0.85\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixb-1} \end{center}

\begin{center}\includegraphics[width=0.87\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixc-1} \end{center}

\begin{center}\includegraphics[width=0.89\linewidth,height=1\textheight]{paperDraft_files/figure-latex/figposteriorPrCheckAppendixd-1} \end{center}

\hypertarget{word-lists}{%
\subsection{Word lists}\label{word-lists}}

\label{appendix:word}

\hypertarget{lists-used-in-previous-research}{%
\subsubsection{Lists used in previous
research}\label{lists-used-in-previous-research}}

\label{appendix:manzini_word_lists} The lists from {[}12{]} are
available here:

\begin{itemize}
\item
  \textbf{Gender:}
  \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/data/vocab/gender_attributes_optm.json}
\item
  \textbf{Race:}
  \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/data/vocab/race_attributes_optm.json}
\item
  \textbf{Religion:}
  \url{https://github.com/TManzini/DebiasMulticlassWordEmbedding/blob/master/Debiasing/data/vocab/religion_attributes_optm.json}
\end{itemize}

The lists are:

\begin{itemize}
\tightlist
\item
  \textbf{Gender:}
\end{itemize}

Pairs of protected attributes: {[}``he'', ``she''{]}, {[}``his'',
``hers''{]}, {[}``son'', ``daughter''{]}, {[}``father'', ``mother''{]},
{[}``male'', ``female''{]}, {[}``boy'', ``girl''{]}, {[}``uncle'',
``aunt''{]}

Lists of stereotypes per gender: man: {[}``manager'', ``executive'',
``doctor'', ``lawyer'', ``programmer'', ``scientist'', ``soldier'',
``supervisor'', ``rancher'', ``janitor'', ``firefighter'',
``officer''{]}, woman: {[}``secretary'', ``nurse'', ``clerk'',
``artist'', ``homemaker'', ``dancer'', ``singer'', ``librarian'',
``maid'', ``hairdresser'', ``stylist'', ``receptionist'',
``counselor''{]} \}

\begin{itemize}
\tightlist
\item
  \textbf{Race:}
\end{itemize}

Sets of protected attributes: {[}``black'', ``caucasian'', ``asian''{]},
{[}``african'', ``caucasian'', ``asian''{]}, {[}``black'', ``white'',
``asian''{]}, {[}``africa'', ``america'', ``asia''{]}, {[}``africa'',
``america'', ``china''{]}, {[}``africa'', ``europe'', ``asia''{]}

Lists of stereotypes per race: ``caucasian'': {[}``manager'',
``executive'', ``redneck'', ``hillbilly'', ``leader'', ``farmer''{]},
``asian'' : {[}``doctor'', ``engineer'', ``laborer'', ``teacher''{]},
``black'' : {[}``slave'', ``musician'', ``runner'', ``criminal'',
``homeless''{]}

\begin{itemize}
\tightlist
\item
  \textbf{Religion:}
\end{itemize}

Sets of protected attributes: {[}``judaism'', ``christianity'',
``islam''{]}, {[}``jew'', ``christian'', ``muslim''{]},
{[}``synagogue'', ``church'', ``mosque''{]}, {[}``torah'', ``bible'',
``quran''{]}, {[}``rabbi'', ``priest'', ``imam''{]}

Lists of stereotypes per race: ``jew'' : {[}``greedy'', ``cheap'',
``hairy'', ``liberal''{]}, ``christian'' : {[}``judgemental'',
``conservative'', ``familial''{]}, ``muslim'' : {[}``violent'',
``terrorist'', ``dirty'', ``uneducated''{]}

\hypertarget{custom-lists-used-in-this-paper}{%
\subsubsection{Custom lists used in this
paper}\label{custom-lists-used-in-this-paper}}

\begin{itemize}
\tightlist
\item
  \textbf{Neutral:}
\end{itemize}

{[}`ballpark', `glitchy', `billy', `dallas', `rip', `called',
`outlooks', `floater', `rattlesnake', `exports', `recursion',
`shortfall', `corrected', `solutions', `diagnostic', `patently',
`flops', `approx', `percents', `lox', `hamburger', `engulfed',
`households', `north', `playtest', `replayability', `glottal',
`parable', `gingers', `anachronism', `organizing', `reach', `shtick',
`eleventh', `cpu', `ranked', `irreversibly', `ponce', `velociraptor',
`defects', `puzzle', `smasher', `northside', `heft', `observation',
`rectum', `mystical', `telltale', `remnants', `inquiry', `indisputable',
`boatload', `lessening', `uselessness', `observes', `fictitious',
`repatriation', `duh', `attic', `schilling', `charges', `chatter',
`pad', `smurfing', `worthiness', `definitive', `neat', `homogenized',
`lexicon', `nationalized', `earpiece', `specializations', `lapse',
`concludes', `weaving', `apprentices', `fri', `militias',
`inscriptions', `gouda', `lift', `laboring', `adaptive', `lecture',
`hogging', `thorne', `fud', `skews', `epistles', `tagging', `crud',
`two', `rebalanced', `payroll', `damned', `approve', `reason',
`formally', `releasing', `muddled', `mineral', `shied', `capital',
`nodded', `escrow', `disconnecting', `marshals', `winamp', `forceful',
`lowes', `sip', `pencils', `stomachs', `goff', `cg', `backyard',
`uprooting', `merging', `helpful', `eid', `trenchcoat', `airlift',
`frothing', `pulls', `volta', `guinness', `viewership', `eruption',
`peeves', `goat', `goofy', `disbanding', `relented', `ratings',
`disputed', `vitamins', `singled', `hydroxide', `telegraphed',
`mercantile', `headache', `muppets', `petal', `arrange', `donovan',
`scrutinized', `spoil', `examiner', `ironed', `maia', `condensation',
`receipt', `solider', `tattooing', `encoded', `compartmentalize',
`lain', `gov', `printers', `hiked', `resentment', `revisionism',
`tavern', `backpacking', `pestering', `acknowledges', `testimonies',
`parlance', `hallucinate', `speeches', `engaging', `solder',
`perceptive', `microbiology', `reconnaissance', `garlic', `neutrals',
`width', `literaly', `guild', `despicable', `dion', `option',
`transistors', `chiropractic', `tattered', `consolidating', `olds',
`garmin', `shift', `granted', `intramural', `allie', `cylinders',
`wishlist', `crank', `wrongly', `workshop', `yesterday', `wooden',
`without', `wheel', `weather', `watch', `version', `usually', `twice',
`tomato', `ticket', `text', `switch', `studio', `stick', `soup',
`sometimes', `signal', `prior', `plant', `photo', `path', `park',
`near', `menu', `latter', `grass', `clock'{]}

\begin{itemize}
\tightlist
\item
  \textbf{Human-related:}
\end{itemize}

{[}`wear', `walk', `visitor', `toy', `tissue', `throw', `talk', `sleep',
`eye', `enjoy', `blogger', `character', `candidate', `breakfast',
`supper', `dinner', `eat', `drink', ``carry'', ``run'', ``cast'',
``ask'', ``awake'', ``ear'', ``nose'', ``lunch'', ``coalition'',
``policies'', ``restaurant'', ``stood'', ``assumed'', ``attend'',
``swimming'', ``trip'', ``door'', ``determine'', ``gets'', ``leg'',
``arrival'', ``translated'', ``eyes'', ``step'', ``whilst'',
``translation'', ``practices'', ``measure'', ``storage'', ``window'',
``journey'', ``interested'', ``tries'', ``suggests'', ``allied'',
``cinema'', ``finding'', ``restoration'', ``expression'',``visitors'',
``tell'', ``visiting'', ``appointment'', ``adults'', ``bringing'',
``camera'', ``deaths'', ``filmed'', ``annually'', ``plane'', ``speak'',
``meetings'', ``arm'', ``speaking'', ``touring'', ``weekend'',
``accept'', ``describe'', ``everyone'', ``ready'', ``recovered'',
``birthday'', ``seeing'', ``steps'', ``indicate'', ``anyone'',
``youtube''{]}

\end{document}