### 2023-10-19 codeTutorial.Rmd

---
title: "R Notebook"
output: html_notebook
---

# Goal

Preparing a Bayesian model for custom data.

# Steps

Load libraries.

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = '../')
library("viridis") 
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(dplyr)
library(dagitty)
library(tidyverse)
library(magrittr)
library(kableExtra)
library(ggpubr)
library(ggExtra)
library(ggthemes)
library(ggforce)
library(rethinking)
library(latex2exp)
library(Hmisc)
library(grid)

```

Load datasets.

```{r cosineDistributions6d,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "75%", warning = FALSE, message = FALSE}

genderGoogle <- read.csv("./datasets/macWeatDatasets/gender_group_google_dataset.csv")[,-1]
genderReddit <-  read.csv("./datasets/macWeatDatasets/gender_group_reddit_dataset.csv")[,-1]
genderGlove <- read.csv("./datasets/macWeatDatasets/gender_group_glove_dataset.csv")[,-1]

raceGlove <-  read.csv("./datasets/macWeatDatasets/race_group_glove_dataset.csv")[,-1]
raceGoogle <- read.csv("./datasets/macWeatDatasets/race_group_google_dataset.csv")[,-1]
raceReddit <- read.csv("./datasets/macWeatDatasets/race_group_reddit_dataset.csv")[,-1]

religionGlove <- read.csv("./datasets/macWeatDatasets/religion_group_glove_dataset.csv")[,-1]
religionGoogle <- read.csv("./datasets/macWeatDatasets/religion_group_google_dataset.csv")[,-1]
religionReddit <- read.csv("./datasets/macWeatDatasets/religion_group_reddit_dataset.csv")[,-1]

debiasedGenderReddit <- read.csv("datasets/macWeatDatasets/debiased_gender_reddit.csv")[-1]
debiasedRaceReddit <- read.csv("datasets/macWeatDatasets/debiased_race_reddit.csv")[-1]
debiasedReligionReddit <- read.csv("datasets/macWeatDatasets/debiased_religion_reddit.csv")[-1]
```

Define a function to clean the data.

```{r}
cleanDataset <- function (dataset) {

colnames(dataset) <- c("pw","word","stereotype", "distance", "similarity", "connection")
dataset$pw <- as.factor(dataset$pw)
dataset$word <- as.factor(dataset$word)
dataset$stereotype <- as.factor(dataset$stereotype)
dataset$connection <- as.factor(dataset$connection)

ifelse(sum( dataset$similarity < -1 | dataset$distance < -1) != 0, 
       print(
         paste("WARNING:  ",  sum( dataset$similarity < -1 | dataset$distance < -1), 
               " out of ", nrow(dataset), " (", sum( dataset$similarity < -1 | dataset$distance < -1)/nrow(dataset) * 100, 
               "%) missing comparisons have been removed!", sep = "")
       ),
       print("No word removal needed.")
)

dataset <- dataset[  dataset$similarity >= -1 | dataset$distance >= -1,]

dataset$associated <- as.integer(dataset$connection == "associated")
dataset$different <- as.integer(dataset$connection == "different")
dataset$human <- as.integer(dataset$connection == "human")
dataset$none <- as.integer(dataset$connection == "none")
dataset$pwi <- as.integer(dataset$pw)
return(dataset)
}
```

Define a function to build models.

```{r}
buildModel <- function(dataset){
options(buildtools.check = function(action) TRUE )
modelResult <- ulam(
  alist(
    distance ~ dnorm(mu,sigma),
    mu <- m +  d[pwi] * different + a[pwi] * associated + h[pwi] * human + n[pw] * none,
    m ~ dnorm(1,.5),
    d[pwi] ~ dnorm(dbar, dsigmabar),
    a[pwi] ~ dnorm(abar, asigmabar),
    h[pwi] ~ dnorm(hbar, hsigmabar),
    n[pwi] ~ dnorm(nbar, nsigmabar),
    dbar ~ dnorm(0,.2),
    abar ~ dnorm(0,.2),
    hbar ~ dnorm(0,.2),
    nbar ~ dnorm(0,.2),
    dsigmabar ~ dexp(5),
    asigmabar ~ dexp(5),
    hsigmabar ~ dexp(5),
    nsigmabar ~ dexp(5),
    sigma ~ dexp(5)
  ),   data = dataset, chains=2 , iter=8000 , warmup=1000,  log_lik = TRUE, cores = 4
)
return(modelResult)
}
```

Define helper functions.

```{r}
savePrecis <- function(modelResult,name){
precisResult <- precis(modelResult, depth = 2)
saveRDS(precisResult,file = paste("resultsDFs/", name, "DF.rds", sep = ""))
}

plotFromPrecis <- function (precis, dataset){

  DFoverall <- rbind(head(precis,1),head(tail(precis, 9),4))
  DFindividual <- head(precis[-1,], -9)

  DFoverall <- DFoverall[-1,]
  DFoverall$type <- c("different", "associated", "human", "none")
  colnames(DFoverall) <- c("mean", "sd", "low", "high", "neff", "rhat", "type")

  ifelse(nrow(DFindividual)/4 != nlevels(dataset$pw),
         print("WARNING: dataset and precis don't match!"),
         print("Dataset and precis match in size"))

  DFindividual$pw <- as.factor(levels(dataset$pw)[rep((1:(nrow(DFindividual)/4)),4)])
  DFindividual$type <- as.factor(c(rep("different", nrow(DFindividual)/4),rep("associated", 
                                                                              nrow(DFindividual)/4),rep("human", nrow(DFindividual)/4), 
                                   rep("none", nrow(DFindividual)/4)))

  pwPlot <- ggplot() + geom_point(data = DFindividual, aes( x = reorder(pw, desc(pw)), y = mean, color = type),
                                  position = position_nudge((as.integer(DFindividual$type)-1)*0.15), size = 1)+
    coord_flip()+theme_tufte()+expand_limits(x= c(-1, nrow(DFindividual)/4 +1))+
    scale_color_manual(values = c("associated" ="orangered4", 
                                  "different" = "chartreuse4",
                                  "human" = "skyblue",
                                  "none" = "grey")) +xlab("protected word")+
    ylab("distance from overall baseline")+
    geom_segment(data = DFindividual,aes(x=reorder(pw, desc(pw)),xend=reorder(pw, desc(pw)), y  = `5.5%`, yend= `94.5%`,
                                         color = type),
                 position = position_nudge((as.integer(DFindividual$type)-1)*0.15), size = .4, alpha = .4)+
    geom_hline(yintercept = 0, lty = 2, size = .2, alpha = .3)+ylim(-.5,.5) +ggtitle("Protected-word-relative coefficients")+ 
    theme(legend.position='none')

  overallPlot <- ggplot() + geom_point(data = DFoverall, aes( x = type, y = mean, color = type), size = 1, position = position_dodge2(width = -1))+
    coord_flip()+theme_tufte() +  
    geom_segment(data = DFoverall,aes(x=type,xend = type, y  = low, yend= high,
                                      color = type))+ylim(-.5,.5)+
    scale_color_manual(values = c("associated" ="orangered4", 
                                  "different" = "chartreuse4",
                                  "human" = "skyblue",
                                  "none" = "grey"))+ theme(legend.position = c(0.2, 0.6))+ggtitle("Overall coefficients")+
    ylab("distance from overall baseline")+
    geom_hline(yintercept = 0, lty = 2, size = .2, alpha = .3)

  plotModel <- grid.arrange(overallPlot,pwPlot, ncol = 1)

  return(plotModel)
}

```

Clean datasets.

```{r}
genderGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_google_dataset.csv")[,-1])
genderReddit <-  cleanDataset(read.csv("./datasets/macWeatDatasets/gender_group_reddit_dataset.csv")[,-1])

raceGlove <-  cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_glove_dataset.csv")[,-1])
raceGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_google_dataset.csv")[,-1])
raceReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/race_group_reddit_dataset.csv")[,-1])

religionGlove <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_glove_dataset.csv")[,-1])
religionGoogle <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_google_dataset.csv")[,-1])
religionReddit <- cleanDataset(read.csv("./datasets/macWeatDatasets/religion_group_reddit_dataset.csv")[,-1])
```

Build exemplary model and save it.

```{r}
modelGenderGoogle <- buildModel(genderGoogle)
saveRDS(modelGenderGoogle, file = "models/modelGenderGoogle.RDS")
```