### 2023-10-19 contextualModelsPOC.Rmd

---
title: "contextualModelsPOC"
output: html_document
date: "2023-10-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Applying Bayesian model for contextual language models.

## List of sentences

When using contextual models, we need to provide the model with a sentence in order to get embeddings for each token depending on a context. Therefore, to perform such a bias analysis on the model's embeddings, we would need a new list of sentences representing the bias and a list of neutral words.

## Limitations

As embedding depends on the context of a word, the use of methods that focus on specific embeddings may be limited. Such lists of phrases to be used should be prepared with care and the limitations should be known and listed in advance.

## Application

Once the lists have been created, the analysis can be carried out in the same way as for the uncontextual models.